---
title: "Increase profits of the bike sharing scheme"
#author: "LKB"
output: 
  pdf_document: 
    fig_caption: no
    highlight: espresso
  html_document: 
    highlight: espresso
    theme: readable
---

```{r, echo=FALSE}
setwd("d:/tmp/Dropbox/Edu/Coursea/StrategicBusinessAnalytics/")
library(knitr)
opts_chunk$set(echo = FALSE, cache = FALSE, cache.path = "cache/", fig.path = "figure/", warning = FALSE,message = FALSE,verbose = FALSE)
#http://yihui.name/knitr/options/
```
Bike sharing schemes are healthy alternative for the car, easing congestion, parking problem and providing much friendlier way to commute. 
```{r}
rm(list=ls(all=TRUE))
data=read.table('day.csv',sep=',',header=TRUE)

data$dteday<-as.Date(data$dteday) #set date correct
data$instant<-NULL
data$yr<-data$yr + 2011

#str(data)
#ummary(data)

usersRatio = sum(data$registered)/sum(data$cnt)
usersRatio = sprintf("%.2f %% of all users",usersRatio*100)

regUsers = aggregate(registered~ workingday, data=data, FUN=sum)
regUsers[2] = regUsers[2]/ sum(data$registered)
regUsersRatio = sprintf("with only %.2f %% active over weekend",regUsers[[2]][1]*100)

casualUsers = aggregate(casual~ workingday, data=data, FUN=sum)
casualUsers[2] = casualUsers[2]/ sum(data$casual)

```
[Data](https://archive.ics.uci.edu/ml/datasets/Bike+Sharing+Dataset) analysis show increased interst in scheme, with registered users constituting `r usersRatio`. This group use bike mostly for work, `r regUsersRatio`.

```{r}
plot(data$dteday,data$cnt,col='red', main = "Bike sharing scheme's rentals",ylab="Number of rentals",xlab = "Date")
points(data$dteday,data$registered,col='blue')
legend("topleft",c("Total rentals","Rentals by casual users"),pch=1,col=c('red','blue'))
```

With a number of casual users visibly increasing during warmer months more targeted campaing is suggested to increase profibility in the biking schemes. To do so further exploration of data is recomended.
```{r}

```

