---
title: "Increase profits of the bike sharing scheme"
#author: "LKB"
classoption: landscape
fontsize: 11pt
#documentclass:
#font-family: 'Courier'
output: 
  pdf_document: 
    fig_caption: no
    #highlight: espresso
   # includes:
    #  in_header: Preamble.tex
---

```{r, echo=FALSE}
setwd("d:/tmp/Dropbox/Edu/Coursea/StrategicBusinessAnalytics/")
library(knitr)
opts_chunk$set(echo = FALSE, cache = FALSE, cache.path = "cache/", fig.path = "figure/", warning = FALSE,message = FALSE,verbose = FALSE,fig.width=12, fig.height=4)
#http://yihui.name/knitr/options/
```

```{r}
rm(list=ls(all=TRUE))
data=read.table('day.csv',sep=',',header=TRUE)

data$dteday<-as.Date(data$dteday) #set date correct
data$instant<-NULL
data$yr<-data$yr + 2011

#str(data)
#ummary(data)
```


```{r}

usersRatio = sum(data$registered)/sum(data$cnt)
usersRatio = sprintf("(%.2f %% of all users)",usersRatio*100)

regUsers = aggregate(registered~ workingday, data=data, FUN=sum)
regUsers[2] = regUsers[2]/ sum(data$registered)
regUsersRatio = sprintf(" (only %.2f %% use)",regUsers[[2]][1]*100)

casualUsers = aggregate(casual~ workingday, data=data, FUN=sum)
casualUsers[2] = casualUsers[2]/ sum(data$casual)
```

Bike sharing schemes offer healthy alternative for the car for urban commuting and sightseeing. It ease congestion, pollution and parking problems. [Existing data](https://archive.ics.uci.edu/ml/datasets/Bike+Sharing+Dataset) analysis show rapidly increasing number of users, though rentals are date dependant. Registered users are the largest segment `r usersRatio`, doing predominantly work committing and not acive during weekends or holidays `r regUsersRatio` . This is partly compensated by casual users being most active during holidays and weekends, especially during warmer months.

```{r}
dev.off() #hack to get std parameters
oldPar<-par() #save std parameters
par(mfrow=c(1,2)) #set plotting to 2 rows
par(mgp=c(1.2,0.4,0), mar=c(2,3,1,1),mai = c(0.5,0.5,0.4,0.5), bty='n')

par(mar=c(5, 4, 4, 2))


plot(data$dteday,data$cnt,col='red', main = "Bike sharing scheme's rentals",ylab="Number of rentals",xlab = "Date")
points(data$dteday,data$registered,col='blue')
legend("topleft",c("Total rentals","Rentals by registered users"),pch=1,col=c('red','blue'))


boxplot(registered~ weekday,data = data, main = "Registered rentals through the week",ylab="Number of rentals",xlab = "Day of week, starting from Sunday") #registerd prefer working days
```

To capitalise on the visible growth I suggest to address both existing and new users with a targeted marketing and promotion campaign, tied up with local event and attractions. We should also confirm durability of infrastructure and maintenance. I suggest: 

* Identify main biking destinations for casual user during holidays/weekends when fleet is underused and tie bike rentals with sightseeing promotions and offers.
* Verify if existing infrastructure is fit for purpose by assessing location of all pick-up stations by analysing patterns of its use and eliminate hot and cold spots;
* Focus on converting local casual users to registered one by promotions;
* Assess feasibility of smaller pick up points in suburbs to advert commuting for casual users;
* Identify main factors for breakage and designing a maintenance regime aiming for shifting bike downtime to the low usage times


```{r}

par(oldPar) #reset plotting parameters

```