---
title: "Strategic Business Analytics, Project preparation"
author: "LKB"
output: 
  html_document: 
    highlight: zenburn
    theme: readable
---


```{r, echo=FALSE}
setwd("d:/tmp/Dropbox/Edu/Coursea/StrategicBusinessAnalytics/")
require(knitr)
opts_chunk$set(echo = TRUE, cache = FALSE, cache.path = "cache/", fig.path = "figure/", warning = FALSE)
#http://yihui.name/knitr/options/
```

##Load data

I plan on working on [Bike Sharing dataset](https://archive.ics.uci.edu/ml/datasets/Bike+Sharing+Dataset).

* instant: record index
* dteday : date
* season : season (1:spring, 2:summer, 3:fall, 4:winter)
* yr : year (0: 2011, 1:2012)
* mnth : month ( 1 to 12)
* hr : hour (0 to 23)
* holiday : weather day is holiday or not (extracted from [Web Link])
* weekday : day of the week
* workingday : if day is neither weekend nor holiday is 1, otherwise is 0.
* weathersit : 
	* 1: Clear, Few clouds, Partly cloudy, Partly cloudy
	* 2: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist
	* 3: Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds
	* 4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog
* temp : Normalized temperature in Celsius. The values are derived via (t-t_min)/(t_max-t_min), t_min=-8, t_max=+39 (only in hourly scale)
* atemp: Normalized feeling temperature in Celsius. The values are derived via (t-t_min)/(t_max-t_min), t_min=-16, t_max=+50 (only in hourly scale)
* hum: Normalized humidity. The values are divided to 100 (max)
* windspeed: Normalized wind speed. The values are divided to 67 (max)
* casual: count of casual users
* registered: count of registered users
* cnt: count of total rental bikes including both casual and registered

```{r, echo=FALSE}
rm(list=ls(all=TRUE))
data=read.table('day.csv',sep=',',header=TRUE)

data$dteday<-as.Date(data$dteday) #set date correct
data$instant<-NULL
data$yr<-data$yr + 2011

str(data)
summary(data)
```

##Exploring data

A quick analysis of everything

```{r}
plot(data$dteday,data$cnt,col='red', main = "Bike users (reg in blue)")
points(data$dteday,data$registered,col='blue')
```
It seems that number of casual users increase only in specific periods of the year.

```{r}
boxplot(cnt ~ season,data = data) #least in spring, most in summer and fall
boxplot(registered ~ season,data = data)
boxplot(casual ~ season,data = data) #least in spring, most in summer and fall
```

And only in specific days.
```{r}
plot(data$dteday,data$registered-data$casual,col='red', main = "Bike users (reg in blue)")
```
While overall number does not change there is a big difference in behaviour of 



```{r}
#much more casual during holidays
boxplot(cnt ~ workingday,data = data) #almost no diff in no, slightly less on working day
boxplot(cnt ~ holiday,data = data) #less users on holidays


boxplot(casual~ weekday,data = data) #casuals prefer holidays
boxplot(registered~ weekday,data = data) #registerd prefer working days

plot(registered~ weathersit,data = data)
windspeed

```


##Lets do some clustering (unsupervised learning)


```{r}
#library(GGally)

plot(atemp ~ temp,data = data) #almost linear correlation use only one

testData<-data
testData <- subset(testData, select = - c(dteday,yr,atemp,mnth,hum,holiday)) #remove columns
#,windspeed,hum))
cor(testData)
#pairs(testData)
#ggpairs(testData,alpha = 0.4 )


testData = scale(testData) 
d = dist(testData, method = "euclidean") # the dist() function computes the distances of all the observations in our dataset
hcward = hclust(d, method="ward.D") # hclust() function performs hiearchical clustering, we pass it the distances, and we set the method argument to "ward.D"
data$groups<-cutree(hcward,k=4) # assign our points to clusters 
```

```{r}
plot(groups ~ hum,data = data) #almost linear correlation use only one
```



```{r}

testdata=data  # To keep our dataset safe, let's create a copy of it called "testdata"

testdata = scale(testdata) 
d = dist(testdata, method = "euclidean") # the dist() function computes the distances of all the observations in our dataset
hcward = hclust(d, method="ward.D") # hclust() function performs hiearchical clustering, we pass it the distances, and we set the method argument to "ward.D"

data$groups<-cutree(hcward,k=3) # assign our points to our k=3 clusters 


```


```{r}


plot(hum ~ cnt,data = data) #almost linear correlation use only one


boxplot(registered ~ season,data = data)
boxplot(casual ~ season,data = data)





boxplot(cnt ~ weathersit,data = data) #most on nice day, almost none on 4+

plot(cnt ~ temp,data = data)

```

* weathersit : 
	* 1: Clear, Few clouds, Partly cloudy, Partly cloudy
	* 2: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist
	* 3: Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds
	* 4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog
	* temp : Normalized temperature in Celsius. The values are derived via (t-t_min)/(t_max-t_min), t_min=-8, t_max=+39 (only in hourly scale)
* atemp: Normalized feeling temperature in Celsius. The values are derived via (t-t_min)/(t_max-t_min), t_min=-16, t_max=+50 (only in hourly scale)
* hum: Normalized humidity. The values are divided to 100 (max)
* windspeed: Normalized wind speed. The values are divided to 67 (max)
* casual: count of casual users

